Quantitative processing of broadband data as implemented in a scientific splitbeam echosounder

### 2022-05-05 Paper/Manuscript.tex
Added rolfs and gavings comments, except figure comments.

The auto correlation function of the matched filter signal and the effective pulse duration, defined as the pulse duration at transmit power $\ptxe$ which produces the same energy as the actual transmitted pulse, will be used in later processing steps and are defined as
\begin{equation}
\label{eq:TXAuto}
\ymfauto(\samplesymt) = \frac{\ymf(\samplesymt)*\ymf^*(-\samplesymt)}{||\ymf||^2_2}
\end{equation}
and
\begin{equation}
\label{eq:TauEff}
\teff = \frac{\sum \ptxauto(\samplesymt)}{\textrm{max}(\ptxauto(\samplesymt))\fsdec},
\end{equation}
where
\begin{equation*}
%\label{eq:TauEff}
\ptxauto(\samplesymt)  =  |\ymfauto(\samplesymt)|^2
\end{equation*}
is the square of the absolute value of the matched filter autocorrelation function, and the summation is calculated over a duration of twice the nominal pulse duration, $2\tnom$. For an ideal system, i.e., no tapering at the rising and trailing edges of the transmitted signal, the effective pulse duration is the same as the transmit pulse duration.

@@

\begin{figure}
\includegraphics[width=16cm]{Fig_TS_echogram}
\caption{\label{Fig_TS_echogram} Sv as a function of ping number and range ($n$) for a calibration sphere. The sphere is located at approximately 6 m range and the seafloor at approximately 12 m. The red vertical line indicates the ping that is used for illustraing the $\ts(\freqsym)$ processing.}
\end{figure}

@@

As a metallic sphere is a rather simple and near-ideal scatterer, we use $\sv$ from a school of non-swimbladdered fish (Fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer to illustrate the processes to obtain $\sv(m)$.

@@

\begin{figure}
\includegraphics[width=16cm]{Fig_Sv_echogram}
\caption{\label{Fig_Sv_echogram} Sv as a function of ping number and range range ($\samplesymt$) for a echogram raw data file. A fish school is seen between 15 and 35 m range, and the sea flor is seen at approximately 50 m. The red vertical line indicates the ping that is used for illustraing the $\sv(\freqsym)$ processing.}
\end{figure}

@@

$\sv(\freqsym)$ is a key parameter for echo integration. To estimate $\sv(\freqsym)$ a Fourier transform is used, repeatedly applied via a sliding window in range. The recommended size of the window is two times the pulse length, and is chosen as a compromise between spatial and frequency resolution. Since the duration of the sliding window can cause the spreading loss compensation to differ between the beginning and end of the window, the compensation for spreading loss is performed on the pulse compressed time domain data before the transform. Absorption loss compensation is also range dependent (and frequency dependent), but is insignificant for typical marine ecosystem echosounder operating frequencies at short ranges and between the beginning and the end of the window at longer ranges. The compensation for absorption loss is therefore performed after applying the discrete Fourier transform. The choice of window also allows the data to be split onto a predefined range-frequency grid, which can then be used to fit data across transducers to an n-dimensional tensor typically employed by machine learning frameworks like Keras, Pytorch, and Tensor flow.

>>>>

The auto correlation function of the matched filter signal and the effective pulse duration, defined as the pulse duration at transmit power $\ptxe$ which produces the same energy as the actual transmitted pulse, will be used in later processing steps and are defined as
\begin{equation}
\label{eq:TXAuto}
\ymfauto(\samplesymt) = \frac{\ymf(\samplesymt)*\ymf^*(-\samplesymt)}{||\ymf||^2_2}
\end{equation}
and
\begin{equation}
\label{eq:TauEff}
\teff = \frac{\sum \ptxauto(\samplesymt)}{\textrm{max}(\ptxauto(\samplesymt))\fsdec},
\end{equation}
where
\begin{equation*}
%\label{eq:TauEff}
\ptxauto(\samplesymt)  =  |\ymfauto(\samplesymt)|^2
\end{equation*}
is the square of the absolute value of the matched filter autocorrelation function, and the summation is calculated over a duration of twice the nominal pulse duration, $2\tnom$. For an ideal system, i.e., no tapering at the rising and trailing edges of the in-water signal, the effective pulse duration is the same as the transmit pulse duration.

@@

\begin{figure}
\includegraphics[width=16cm]{Fig_TS_echogram}
\caption{\label{Fig_TS_echogram} Sv as a function of ping number and range ($n$) for a calibration sphere. The sphere is located at approximately 6 m range. The red vertical line indicates the ping that is used for illustraing the $\ts(\freqsym)$ processing.}
\end{figure}

@@

As a metallic sphere is a rather simple and ideal scatterer suitable for investigating TS of single scatterers, we use $\sv$ from a school of non-swimbladdered fish (Fig.~\ref{Fig_Sv_echogram}) collected with a 120 kHz centre frequency transducer to illustrate the processes to obtain $\sv(m)$.

@@

\begin{figure}
\includegraphics[width=16cm]{Fig_Sv_echogram}
\caption{\label{Fig_Sv_echogram} Sv as a function of ping number and range range ($\samplesymt$) for a echogram raw data file. The upper yellow area shows the transmit pulse, a fish school is seen as a registration between 15 and 35 m range, and the sea flor is seen at approximately 50 m. The red vertical line indicates the ping that is used for illustraing the $\sv(\freqsym)$ processing.}
\end{figure}

@@

$\sv(\freqsym)$ is a key parameter for echo integration. To estimate $\sv(\freqsym)$ a Fourier transform is used, repeatedly applied via a sliding window in range. The recommended size of the window is two times the pulse length, and is chosen as a compromise between spatial and frequency resolution. Since the duration of the sliding window can cause the spreading loss compensation to differ between the beginning and end of the window, the compensation for spreading loss is performed on the pulse compressed time domain data before the transform. Absorption loss compensation is also range dependent (and frequency dependent), but is insignificant for typical marine ecosystem echosounder operating frequencies at short ranges and between the beginning and the end of the window at longer ranges. The compensation for absorption loss is therefore performed after applying the discrete Fourier transform. The choice of window also allows the data to be split onto a predefined range-frequency grid, which can then be used to fit data across transducers to an n-dimensional tensor typically employed by deep learning methods \citep[e.g.]{brautaset_acoustic_2020}.
